---
title: "Week 05 - Homework"
author: "Jose Luis Rodriguez Gil"
date: "14/10/2020"
output: 
  html_document:
    keep_md: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, include = FALSE}

library(tidyverse)
library(janitor)
library(here)

```

## 1. Lad data

We are going to load the required dataset of water quality measured in ditches. 

```{r load data}

dat_original <- read_csv(here("data", "ditch.csv"))

```
# 2. Data processing
 
 Let's clean the names so they fit within the tydiverse style guide
 
```{r data processing}

dat <- dat_original %>% 
  clean_names()

```
 
# 3 Data summary
 
```{r data summary - Item 1}

sulphate_summary <- dat %>% 
  pivot_longer(cols = c(-site, -year, -month, -depth), names_to = "parameter", values_to = "measure") %>% 
  filter(parameter == "sulphate") %>% 
  group_by(site, parameter) %>% 
  summarise(mean_measure = mean(measure, na.rm = TRUE),
            sd_measure = sd(measure, na.rm = TRUE)) %>% 
  mutate(mean_measure = round(mean_measure, digits = 2),
         sd_measure = round(sd_measure, digits = 2))

```

This following chunk is not required for the assignment, but it formats the table so it looked pretty in the `README.md`

```{r, echo = FALSE, results='asis'}

knitr::kable(sulphate_summary)

```


# 4. plots

## 4.1 Bar chart

We will use the generated summary table to create the following bar chart

```{r item_2}

sulphate_summary %>% 
  ggplot() +
  geom_col(aes(x = site, y = mean_measure, fill = site)) +
  geom_errorbar(aes(x = site, ymin = mean_measure - sd_measure, ymax = mean_measure + sd_measure)) +
  labs(x = NULL,
       y = "Sulphate concentration")


```

##4.2 Box plot
 
 Now lets look at the chloride concentrations between sites
 
```{r item_3}

dat %>% 
  pivot_longer(cols = c(-site, -year, -month, -depth), names_to = "parameter", values_to = "measure") %>% 
  filter(parameter == "chloride") %>% 
  ggplot() +
  geom_boxplot(aes(x = site, y = measure, colour = site)) +
  geom_jitter(aes(x = site, y = measure, colour = site), width = 0.1, alpha = 0.4) +
  labs(x = NULL,
       y = "Chloride concentration")

```

##4.3 temporal points plot
 
```{r item_4}

dat %>% 
  pivot_longer(cols = c(-site, -year, -month, -depth), names_to = "parameter", values_to = "measure") %>% 
  filter(parameter == "chloride") %>% 
  ggplot() +
  facet_wrap(~site) +
  geom_point(aes(x = year, y = measure, colour = site)) +
  geom_smooth(aes(x = year, y = measure, colour = site), method = "lm") +
  labs(x = NULL,
       y = "Chloride concentration")

```
 
 

